%-*-LaTeX-*-
\documentclass[11pt]{report}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Use the approved methods for setting lengths
%\oddsidemargin  0.0in
%\evensidemargin 0.0in
%\textwidth      6.5in
%\textheight     9.0in
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.7in}
% Based on document style, and taller text body height, set
% weird LaTex margin/header (box heights).  This fixes
% the disappearing page numbers (which never disappeared; they
% just got printed beyond the borders of the physical paper)
\setlength{\voffset}{0pt}
\setlength{\topmargin}{0pt}
\setlength{\headheight}{10pt}
\setlength{\headsep}{25pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{html}
\usepackage{makeidx}
\usepackage{times}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{graphicx}
%\usepackage{eufrak}
\usepackage{amsmath}

%\RequirePackage{times}
% tux230 doesn't like the index stuff
\makeindex

\tolerance=600

\newcommand{\Fb}{{\bf F}}
\newcommand{\Lb}{{\bf L}}
\newcommand{\gb}{{\bf g}}
\newcommand{\nb}{{\bf n}}
\newcommand{\ub}{{\bf u}}
\newcommand{\xb}{{\bf x}}
\newcommand{\p}{\partial}
\renewcommand{\div}{{\rm div}}
\renewcommand{\arraystretch}{1.3}

\begin{document}

\title{\Huge User's guide to S4W and S4Wopt version 0.0.1} 

\author{ N. Anders Petersson$^1$ \and Bj\"orn Sj\"ogreen\thanks{Center for Applied Scientific
     Computing, Lawrence Livermore National Laboratory, PO Box 808, Livermore CA 94551. This is
     contribution LLNL-SM-487431.}}
\date{\today} 
\maketitle
\pagestyle{myheadings}

%\thispagestyle{plain}
\markboth{}{N.A. PETERSSON AND B. SJOGREEN; USER'S GUIDE TO S4W/S4WOPT-0.0.1}

\pagebreak
\paragraph {Disclaimer} 
This document was prepared as an account of work sponsored by an agency of the United States
government. Neither the United States government nor Lawrence Livermore National Security, LLC, nor
any of their employees makes any warranty, expressed or implied, or assumes any legal liability or
responsibility for the accuracy, completeness, or usefulness of any information, apparatus, product,
or process disclosed, or represents that its use would not infringe privately owned
rights. Reference herein to any specific commercial product, process, or service by trade name,
trademark, manufacturer, or otherwise does not necessarily constitute or imply its endorsement,
recommendation, or favoring by the United States government or Lawrence Livermore National Security,
LLC. The views and opinions of authors expressed herein do not necessarily state or reflect those of
the United States government or Lawrence Livermore National Security, LLC, and shall not be used for
advertising or product endorsement purposes. 

\paragraph{Auspices Statement}
This work performed under the auspices of the U.S. Department of Energy by Lawrence Livermore
National Laboratory under contract DE-AC52-07NA27344.
%\pagebreak
\tableofcontents

% Introduction
\chapter{Source inversion}
\emph{S4Wopt} is the main program for source inversion. \emph{S4Wopt} estimates a source of the form
$$
 f({\bf x},t) =  g(t,t_0,\omega_0)M\nabla\delta({\bf x}-{\bf x}^*)
$$
where, ${\bf x}^*$ is the location of the source, $M$ is the symmetric 3$\times$3 moment tensor, and the time
function $g$ contains two parameters, the frequency, $\omgea_0$, and time shift, $t_0$. Currently, the only
implemented time functions for source inversion are the Gaussian and the VerySmoothBump.
\par
The source is computed by minimizing the misfit between the computed time series ${\bf u}^n_{{\bf i}_r}$ and measured 
data ${\bf d}^n_{{\bf i}_r}}$,
$$
 \chi = \frac{1}{2}\sum_{n=0}^N \sum_{r=1}^R |{\bf u}^n_{{\bf i}_r}-{\bf d}^n_{{\bf i}_r}}|^2,
$$
with respect to the eleven parameters describing the source, ${\bf x}^*, M, \omeaga_0$, and $t_0$.
The measured data are given at stations ${\bf x}_{{\bf i}_r}$, $r=1,\ldots,R$.
The {\tt observation} keyword is used to specify the input seismograms (${\bf d}^n_{{\bf i}_r}$ above). 
For example, the command
%
\begin{verbatim}
observation x=11200 y=6300 z=0.0 file=cartstation
\end{varbatim}
%
will read a file named {\tt cartstation.txt} from the directory specified by the {\tt path} parameter to the {\tt fileio}
command. If the {\tt path} is not set, the file will be searched for in the current directory. Currently, only files on
the USGS ascii format can be read by the observation command. The USGS file do not contain information about the
location of the station where the data were obtained. It is thus necessary to specify the location directly
in the {\tt observation} command. In the example above, the data on the file represent measurements at a 
station located at the position ${\bf x} =  (11200,6300,0)$ in the Cartesian grid. It is possible to specify the
location in geographic coordinates instead. For example,
%
\begin{verbatim}
observation lon=-122.523 lat=38.0496 depth=0.0 file=nhf 
\end{varbatim}
%
will read a file named {\tt nhf.txt}, that represents data obtained at latitude 38.049 and longitude -122.523 on 
the surface of the earth.
\emph{S4Wopt} will interpret the components of read file as Cartesian components, or as components pointing in the 
east, north, and up directions, depending on information in the header of the file.
\par
The file read by the observation command will be assumed to start at time = 0, even if the starting time recorded 
in the file is different from zero, i.e., the read seismograms will be shifted in time to always start at the first
time step of the simulation.
\par
\begin{narrow}
{\emph Not Yet Implemented:}
Plan for reading sac files:
\begin{verbatim}
observation sfile1=filename.x sfile2=filename.y sfile3=filename.z
\end{verbatim}
will read the three components of a time history at a point from three SAC files. Information about type of
components and location of the station is taken from the header of the SAC file.
\end{narrow}
\par
The stations specified by the observation command, will also act as recievers of time series during the 
forward solve phase. After each iteration in the optimization algorithm, the time histories obtained at the stations 
(${\bf u}^n_{{\bf i}_r}$ above) will be output with on files with a {\tt \_out} appended to their names. For example,
if time history is input on the file {\tt nhf.txt}, the corresponding seismograms computed by the forward problem
using the most recent approximation of the source are written on a file named {\tt nhf_out.txt}. These files are thus
overwitten after each optimization iteration. Furthermore, if the option output_initial_seismograms is set to
one, the time series obtained with the initial guess source will be output on files with a {\tt \_ini} appended.
\par
The {\tt receiver} command, defined for \emph{S4W}, can not be used with \emph{S4Wopt}.  Currently, it is not possible to
specify a receiver that is not also an observation.

%\chapter{Introduction}
%\chapter{Getting started}
\chapter{Coordinate system, units and the grid}

The computation takes place over a time interval, $0\leq t \leq T$, where $T$ is a user specified end time.
The starting point of the simulation time is always $t=0$. The {\tt time} command specifies the end time, for example
\begin{verbatim}
time t=1.6
\end{verbatim}
sets $T=1.6$. The unit is seconds. Alternatively, the simulation time interval can be specified as 
a number of time steps, as in the example
\begin{verbatim}
time steps=1200
\end{verbatim}
which sets the number of time steps to 1200. The end time will in this case be $T=1200\Delta t$, where the
time step $\Delta t$ depends on the input material. $\Delta t$ is determined automatically by \emph{sw4}.
The simulation start time can be related to a UTC time. The option {\tt utcstart} sets the UTC time that
corresponds to the simulation time $t=0$, for example,
\begin{verbatim}
time t=1.6 utcstart=01/04/2012:17:34:12.233
\end{verbatim}
The format of the UTC time is ``month/day/year:hour:minute:second.milisecond''. When the UTC time is set,
output receiver data files are time stamped with the UTC reference time. Furthermore, the UTC start time
is essential to align UTC stamped observed data correctly when solving the inverse problem by \emph{sw4opt}.


%\chapter{Sources, time-functions and grid sizes}
%\chapter{The material model}
%\chapter{Topography} \label{sec:topography}
%\chapter{Mesh refinement} \label{sec:mesh-ref}
%\chapter{Attenuation}\label{sec:attenuation}
%\chapter{Output options}
%\chapter{Examples} \label{sec:examples}
%\chapter{Keywords in the input file}
%\chapter{File formats}
%\chapter{Installing \emph{WPP}}\label{cha:installing-wpp}
%\chapter{Testing the \emph{WPP} installation}

\chapter{Keywords in the input file}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The syntax of the input file is
\begin{verbatim}
command1 parameter1=value1 parameter2=value2 ... parameterN=valueN
# comments are disregarded
command2 parameter1=value1 parameter2=value2 ... parameterN=valueN
...
\end{verbatim}
Each command starts at the beginning of the line and ends at the end of the same line. Blank and
comment lines are disregarded. A comment is a line starting with a \# character. The order of the
parameters within each command is arbitrary. The material commands (block, ifile, pfile, and efile)
are applied in the order they appear, but the ordering of all other commands is
inconsequential. Also note that the entire input file is read before the simulation starts.

Parameter values are either integers (-2,0,5,...), floating point numbers (20.5, -0.05, 3.4e4), or strings
(wpp, earthquake, my-favorite-simulation). Note that there must be no spaces around the = signs and
strings are given without quotation marks and must not contain spaces. Depending on the specific
command, some parameter values are required to fall within specified ranges.

A breif description of all commands is given in the following sections. The commands marked as
[required] must be present in all S4W input files, while those marked as [optional] are just
that. Other commands, such as those specifying the material model can be given by (a combination of)
different commands (block, pfile, efile, or ifile). Unless \emph{S4W} is run in one of its test
modes, the material must be specifed by at least one of these commands and at least one source must
be specified.


