%-*-LaTeX-*-
\documentclass[11pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Use the approved methods for setting lengths
%\oddsidemargin  0.0in
%\evensidemargin 0.0in
%\textwidth      6.5in
%\textheight     9.0in
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.7in}
% Based on document style, and taller text body height, set
% weird LaTex margin/header (box heights).  This fixes
% the disappearing page numbers (which never disappeared; they
% just got printed beyond the borders of the physical paper)
\setlength{\voffset}{0pt}
\setlength{\topmargin}{0pt}
\setlength{\headheight}{10pt}
\setlength{\headsep}{25pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{color}
\usepackage{textpos}
%\usepackage{html}
\usepackage{makeidx}
%\usepackage{times}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{graphicx}
%\usepackage{eufrak}
\usepackage{amsmath}

\makeindex

\tolerance=600

\newcommand{\Fb}{{\bf F}}
\newcommand{\Lb}{{\bf L}}
\newcommand{\gb}{{\bf g}}
\newcommand{\nb}{{\bf n}}
\newcommand{\ub}{{\bf u}}
\newcommand{\xb}{{\bf x}}
\newcommand{\p}{\partial}
\renewcommand{\div}{{\rm div}}
\renewcommand{\arraystretch}{1.3}

\begin{document}

\title{Installing SW4 version 3.0}

\author{ N. Anders Petersson$^*$ \and Bj\"orn Sj\"ogreen\thanks{Center for Applied Scientific
    Computing, Lawrence Livermore National Laboratory, PO Box 808, Livermore CA 94551. This work
    performed under the auspices of the U.S. Department of Energy by Lawrence Livermore National
    Laboratory under contract DE-AC52-07NA27344. This is contribution
    LLNL-SM-741310.} 
\and Houjun Tang\thanks{ Lawrence Berkeley National Laboratory}
    }
\date{September 21, 2022}
\maketitle

%\thispagestyle{plain}

%\pagebreak


%\pagebreak
\tableofcontents

\section{Introduction}
The sole purpose of this document is to describe the installation process of the seismic wave
propagation code \emph{SW4}. A comprehensive user's guide is provided in the report by
Petersson and Sjogreen~\cite{SW4-v3}.

\section{Compilers and third party libraries}

Before you can build \emph{SW4} on your system, you must have
\begin{enumerate}
\item the \verb+lapack+ and \verb+blas+ libraries. These libraries provide basic linear algebra functionality and are pre-installed on many machines;
\item an MPI library. This library provides support for message passing on parallel
  machines. Examples of open source implementations include Mpich and OpenMPI. Note that the MPI library must be installed even if you are only building \emph{SW4} for a single core system.
\end{enumerate}

To avoid incompatibility issues and linking problems, we recommend using the same compiler for the libraries as for \emph{SW4}.

In order to use geographic projection and material models stored in the \emph{rfile} format, you need to install the PROJ before building \emph{SW4}:
\begin{itemize}
\item PROJ (version 6+), \url{https://proj.org/download.html}
\end{itemize}  
If you also wish to use material models using the \emph{sfile} or \emph{GMG} format, you also need to download and install the HDF5 library:
\begin{itemize}
\item HDF5 (version 1.12+), \url{https://www.hdfgroup.org/downloads/hdf5/source-code}
\end{itemize}
To simplify the build process, all libraries should be installed under the same directory, such that the library files (.so, .a, etc.) are in the lib sub-directory and the include files (.h) end up in the include sub-directory. See Section~\ref{sec:proj} for details.

\paragraph{MacOS}

We recommend using the MacPorts package manager for installing the required compilers and libraries. Simply go to www.macports.org, and install macports on your system. With that in place, you can use the \verb+port+ command as follows
\begin{verbatim}
sudo port install gcc11
sudo port install mpich-gcc11
sudo port install hdf5
\end{verbatim}
Here, \verb+gcc11+ refers to version 11 of the GNU compiler suite. Compiler versions are bound to change in the future, so the above commands will need to be modified accordingly. Before starting, make sure you install a version of gcc that is compatible with the MPI library package. The above example installs the \verb+mpich+ package using the gcc11 compilers, which includes a compatible Fortran compiler. Alternatively, you can use the \verb+openmpi+ package. Note that the \verb+port select+ commands are used to create shortcuts to the compilers and MPI environment. By using the above setup, the GNU compilers can be accessed with \verb+gcc+ and \verb+gfortran+ commands, and the MPI compilers and execution environment are called \verb+mpicxx+, \verb+mpif90+, and \verb+mpirun+, respectively.

The \verb+lapack+ and \verb+blas+ libraries are preinstalled on recent Macs and can be accessed using the \verb+-framework Accelerate+ link option. If that is not available or does not work on your machine, you can download \verb+lapack+ and \verb+blas+ from \url{www.netlib.org.}

\paragraph{Linux}
We here give detailed instructions for installing the third part libraries under 64 bit, Fedora Core 18 Linux. Other Linux variants use similar commands for installing software packages, but note that the package manager \verb+yum+ is specific to Fedora Core. For Ubuntu systems, one can simply change \verb+yum+ to \verb+apt-get+

You need to have root privileges to install precompiled packages. Start by opening an xterm and set your user identity to root by the command
\begin{verbatim}
su -
\end{verbatim}
Install the compilers by issuing the commands
\begin{verbatim}
yum install gcc
yum install gcc-c++
yum install gcc-gfortran
\end{verbatim}
You install the \verb+mpich+ and \verb+hdf5+ library and include files with the command
\begin{verbatim}
yum install mpich-devel
yum install hdf5
\end{verbatim}

The executables and libraries are installed in \verb+/usr/lib64/mpich/bin+ and  \verb+/usr/lib64/mpich/lib+ respectively. We suggest that you add \verb+/usr/lib64/mpich/bin+ to your path. This is done with the command:
\begin{verbatim}
export PATH=${PATH}:/usr/lib64/mpich/bin
\end{verbatim}
if your shell is \verb+bash+. For \verb+tcsh+ users, the command is
\begin{verbatim}
setenv PATH ${PATH}:/usr/lib64/mpich/bin
\end{verbatim}
It is convenient to put the path setting command in your startup file, \verb+.bashrc+ or \verb+.cshrc+., for bash or csh/tcsh respectively.
The \verb+blas+ and \verb+lapack+ libraries are installed with
\begin{verbatim}
yum install blas
yum install lapack
\end{verbatim}
On our system, the libraries were installed in \verb+/usr/lib64+ as \verb+libblas.so.3+ and \verb+liblapack.so.3+. For some unknown reason, the install program does not add links to these files with extension \verb+.so+, which is necessary for the linker to find them. We must therefore add the links explicitly. If the libraries were installed elsewhere on your system, but you don't know where, you can find them with the following command:
\begin{verbatim}
find / -name "*blas*" -print
\end{verbatim}
After locating the directory where the libraries reside (in this case \verb+/usr/lib64+), we add
links to the libraries with the commands:
\begin{verbatim}
cd /usr/lib64
ln -s libblas.so.3 libblas.so
ln -s liblapack.so.3 liblapack.so
\end{verbatim}
Note that you need to have \verb+root+ privileges for this to work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Unpacking the source code tar ball}
\index{installation!directories}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To unpack the \emph{SW4} source code, you place the file \verb+sw4-v3.0.tar.gz+ in the
desired directory and issue the following command:
\begin{verbatim}
tar xzf sw4-v3.0.tar.gz
\end{verbatim}
As a result a new sub-directory named \verb+sw4-v3.0+ is created. It contains several files
and sub-directories:
%
\begin{itemize}
\item \verb+LICENSE.txt+ License information.
\item \verb+INSTALL.txt+ A link to this document.
\item \verb+README.txt+ General information about \emph{SW4}.
\item \verb+configs+ Directory containing \verb+make+ configuration files.
\item \verb+src+ C++ and Fortran source code of \emph{SW4}.
\item \verb+tools+ Matlab/Octave scripts for post processing and analysis.
\item \verb+pytest+ Python script and input files for testing the \emph{SW4} installation.
\item \verb+examples+ Sample input files.
\item \verb+Makefile+ Main makefile (don't change this file!).
\item \verb+CMakeLists.txt+ CMake configuration file (don't change this file either!).
\item \verb+wave.txt+ Text for printing the "SW4 Lives" banner at the end of a successful build.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Installing \emph{SW4} with make}\label{cha:installing-sw4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%The \emph{SW4} source code is released under the GNU general public license and can be downloaded
%from:
%\begin{verbatim}
%https://computation.llnl.gov/casc/serpentine/software.html
%\end{verbatim}

The classical way of building \emph{SW4} uses \verb+make+. We recommend using GNU make, sometimes called \verb+gmake+. You can check the version of make on you system with the command
\begin{verbatim}
make -v
\end{verbatim}
If you don't have GNU make installed on your system, you can obtain it from www.gnu.org.

We have built \emph{SW4} and its supporting libraries on Intel based laptops and desktops running
LINUX and OSX. It has also been built on several supercomputers such as the Intel machines {\tt 
quartz} (at LLNL), {\tt Cori} (at LBNL), and {\tt Summit} (at ORNL). We have
successfully used the following versions of Gnu, Intel, and IBM compilers:
\begin{verbatim}
GNU:   g++/gcc/gfortran      versions 4.5+
Intel: icpc/icc/ifort        versions 16.0+
IBM:   xlcxx/xlc/xlf         versions 12.1+
\end{verbatim}

\emph{SW4} uses the message passing interface (MPI) standard for communication on parallel distributed memory machines. Note that the MPI library often includes wrappers for compiling, linking, and running of MPI programs. For example, the {\tt mpich} package build wrappers for the underlying C++ and Fortran compilers called {\tt mpicxx} and {\tt mpif90}, as well as the {\tt mpirun} script. We highly recommend using these programs for compiling, linking, and running \emph{SW4}.


\subsection{Basic compilation and linking of \emph{SW4}}\label{sec:basic-install}
\index{installation!basic} 

The basic build process is controlled by the environmental variables \verb+FC+, \verb+CXX+, \verb+EXTRA_FORT_FLAGS+, \verb+EXTRA_CXX_FLAGS+, and \verb+EXTRA_LINK_FLAGS+. These variables should hold the names of the Fortran and C++ compilers, and any extra options that should be passed to the compilers and linker. The easiest way of assigning these variables is by creating a file in the \verb+configs+ directory called \verb+make.inc+. The \verb+Makefile+ will look for this file and read it if it is available. There are several examples in the \verb+configs+ directory, e.g. \verb+make.osx+ for Macs and \verb+make.linux+ for Linux machines. You should copy one of these files to your own \verb+make.inc+ and edit it as needed. 

\subsubsection{Mac machines}
If you are on a Mac, you could copy the setup from \verb+make.osx+,
\begin{verbatim}
cd configs
cp make.osx make.inc
cat make.inc
proj = no
hdf5 = no
FC = mpif90
CXX = mpicxx
EXTRA_FORT_FLAGS = 
EXTRA_LINK_FLAGS = -framework Accelerate -L/opt/local/lib/gcc11 -lgfortran
\end{verbatim}
In this case, the \verb+blas+ and \verb+lapack+ libraries are assumed to be provided by the \verb+-framework Accelerate+ option. The \verb+libgfortran+ library is located in the directory \verb+/opt/local/lib/gcc11+, which is where \verb+macports+ currently installs it.

\subsubsection{Linux machines}
If you are on a Linux machine, we suggest you copy the configuration options from \verb+make.linux+,
\begin{verbatim}
cd configs
cp make.linux make.inc
cat make.inc
proj = no
FC = gfortran
CXX = mpicxx
EXTRA_LINK_FLAGS = -L/usr/lib64 -llapack -lblas -lgfortran
\end{verbatim}
This setup assumes that the \verb+blas+ and \verb+lapack+ libraries are located in the
\verb+/usr/lib64+ directory. 
In the case of Fedora Core 18, we needed to set the link flag variable to
\begin{verbatim}
EXTRA_LINK_FLAGS = -Wl,-rpath=/usr/lib64/mpich/lib  -llapack -lblas -lgfortran
\end{verbatim}

\subsubsection{Using make}
You build \emph{SW4} with the "make" command from the main directory.
\begin{verbatim}
cd /enter/your/path/sw4-v3.0
make
\end{verbatim}
If all goes well, you will see the SW4 Lives banner on your screen after the compilation and linking
has completed,
{\samepage
\begin{verbatim}
    
``'-.,_,.-'``'-.,_,.='``'-.,_,.-'``'-.,_,.='````'-.,_,.-'``'-.,_,.='``
  _________    ____      __      ____    _    __
 /   ____  \   \   \    /  \    /   /   / |  |  |  
 |  |    \./    \   \  /    \  /   /   /  |  |  |
 |  |______      \   \/      \/   /   /   '--'  |
 \______   \      \              /    |______   |
        |  |       \     /\     /            |  |      
 /`\____|  |        \   /  \   /             |  |      
 \_________/         \_/    \_/              |__|      
                                       
   __       __  ____    ____  _______    ______    __  
  |  |     |  | \   \  /   / |   ____|  /    __|  |  | 
  |  |     |  |  \   \/   /  |  |__     |   (__   |  | 
  |  |     |  |   \      /   |   __|    \__    |  |  | 
  |  `----.|  |    \    /    |  |____    __)   |  |__| 
  |_______||__|     \__/     |_______|  (_____/   (__)
``'-.,_,.-'``'-.,_,.='``'-.,_,.-'``'-.,_,.='````'-.,_,.-'``'-.,_,.='``

\end{verbatim}
}
By default, \verb+make+ builds an optimized \verb+sw4+ executable. It is located in 
\begin{verbatim}
/enter/your/path/sw4-v3.0/optimize/sw4
\end{verbatim}
You can also build an executable with debugging symbols by adding the \verb+debug=yes+ option to \verb+make+,
\begin{verbatim}
cd /enter/your/path/sw4-v3.0
make debug=yes
\end{verbatim}
In this case, the executable will be located in
\begin{verbatim}
/enter/your/path/sw4-v3.0/debug/sw4
\end{verbatim}
It can be convenient to add the corresponding directory to your \verb+PATH+ environment
variable. This can be accomplished by modifying your shell configuration file, e.g. \verb+~/.cshrc+
if you are using C-shell.

\subsubsection{How do I setup the {\tt make.inc} file?}

The input file for \verb+make+ is
\begin{verbatim}
sw4-v3.0/Makefile
\end{verbatim}
Do {\em not} change this \verb+Makefile+. It should only be necessary to edit your configuration
file, that is,
\begin{verbatim}
/my/path/sw4-v3.0/configs/make.inc
\end{verbatim}
Note that you must create this file, for example by copying one of the \verb+make.xyz+ files in the
same directory. The \verb+make.inc+ file holds all information that is particular for your system,
such as the name of the compilers, the location of the third party libraries, and any extra
arguments that should be passed to the compiler or linker. 

The following \verb+make.inc+ file includes all configurable options:
\begin{verbatim}
proj = no
hdf5 = no
SW4ROOT = /my/path/to/installed/library
CXX = mpicxx
FC  = mpif77
EXTRA_CXX_FLAGS = -DUSING_MPI
EXTRA_FORT_FLAGS = -fno-underscoring
EXTRA_LINK_FLAGS = -framework vecLib
\end{verbatim}
The \verb+CXX+ and \verb+FC+ variables should be set to the names of the C++ and
Fortran compilers, respectively. Finally, the \verb+EXTRA_CXX_FLAGS+, \verb+EXTRA_FORT_FLAGS+, and
\verb+EXTRA_LINK_FLAGS+ variables should contain any additional arguments that need to be passed to
the C++ compiler, Fortran compiler, or linker, on your system.

\subsection{Building \emph{SW4} with PROJ, HDF5, and ZFP support}
The PROJ libraray enables the more advanced geographical mapping keywords in the {\tt grid} command and is also required by the {\tt rfile, sfile} and {\tt gmg} commands. To enable the {\tt sfile}, {\tt ssioutput}, and {\tt gmg} commands, you have to also install the {\tt HDF5} library. To use ZFP compression for the {\tt ssioutput} command, {\tt ZFP} and {\tt H5Z-ZFP} are required. See Section \ref{sec:proj} for installing these libraries.

Once you have successfully installed the PROJ, and optionally the HDF5 and ZFP libraries, it should be easy to re-configure \emph{SW4} to use them. Simply edit your configuration file (\verb+make.inc+) by adding the following lines to the top of the file, setting the {\tt proj}, {\tt hdf5}, and {\tt zfp} keywords to
{\tt yes} or {\tt no}, as appropriate.
\begin{verbatim}
proj = yes 
hdf5 = yes
zfp = yes
SW4ROOT = /thid/party/basedir
HDF5ROOT = /thid/party/basedir
ZFPROOT = /thid/party/basedir
H5ZROOT = /thid/party/basedir
\end{verbatim}

You then need to re-compile \emph{SW4}. Go to the \emph{SW4} main directory, clean out the previous
object files and executable, and re-run make:
\begin{verbatim}
make clean
make
\end{verbatim}
If all goes well, the ``SW4 lives'' banner is shown after the make command is
completed. As before, the \verb+sw4+ executable will be located in the \verb+optimize+ or
\verb+debug+ directories.

\subsection{Testing the \emph{SW4} installation}
The \emph{SW4} source code distribution includes a python(3) script for running several tests and
checking the solutions against previously verified results. Note that the same set of tests can be
performed when \emph{SW4} is built with CMake, see Section~\ref{cha:ctest-sw4}.

After \emph{SW4} has been built with \verb+make+, go to the \verb+pytest+ directory and run
\verb+test_sw4.py+. If the \verb+sw4+ executable resides in the \verb+optimize+ directory, you can
run the basic tests by doing:
\begin{verbatim}
cd pytest
./test_sw4.py ("./test_sw4.py -u 0" if HDF5 is not installed)
\end{verbatim}
If all goes well, you should see the following output:
\begin{verbatim}
Running all tests for level 0 ...
Test # 1 Input file: energy-nomr-2nd-1.in PASSED
Test # 2 Input file: energy-mr-4th-1.in PASSED
...
Test # 25 Input file: loh1-h100-mr-restart-hdf5-1.in PASSED
Out of 25 tests, 0 failed, 23 passed, and 2 skipped
\end{verbatim}
Some aspects of the testing can be modified by providing command line arguments to
\verb+test_sw4.py+. For a complete list of options do \verb+test_sw4.py --help+, which currently
give the output:
\begin{verbatim}
./test_sw4.py --help
usage: test_sw4.py [-h] [-v] [-l {0,1,2}] [-m MPITASKS] [-d SW4_EXE_DIR]

optional arguments:
  -h, --help            show this help message and exit
  -v, --verbose         increase output verbosity
  -l {0,1,2}, --level {0,1,2}
                        testing level
  -m MPITASKS, --mpitasks MPITASKS
                        number of mpi tasks
  -t OMPTHREADS, --ompthreads OMPTHREADS
                        number of omp threads per task
  -d SW4_EXE_DIR, --sw4_exe_dir SW4_EXE_DIR
                        name of directory for sw4 executable
  -p PYTEST_DIR, --pytest_dir PYTEST_DIR
                        full path to the directory of pytest
                        (/path/sw4/pytest)
  -u {0,1,2,3,4}, --usehdf5 {0,1,2,3,4}
                        run HDF5 tests with, 0 no HDF5, 1 first HDF5 case, 2
                        second case,..., 4 all cases
  -g {0,1}, --geodynbc {0,1}
                        run Geodynbc tests with, 0 skip, 1 both cases
  -A CPU_ALLOCATION, --cpu_allocation CPU_ALLOCATION
                        name of cpu bank/allocation
\end{verbatim}
Note  that the directory name for the \verb+sw4+ executable should be given relative to the main
\verb+sw4+ directory.

\section{Installing \emph{SW4} with CMake}\label{cha:installing-cmake-sw4}
\emph{SW4} can also be built with CMake. Compared to using regular {\tt make}, this build process is
easier to use because it is fully automated. However, it gives the user less control of which
compilers, linker, and libraries to use. Similar to using regular {\tt make}, the \emph{SW4} CMake
configuration allows automated correctness testing of the installation. The test runs the same set
of cases as the \verb+test_sw4.py+ script in the \verb+pytest+ directory, see
Section~\ref{cha:ctest-sw4} for details.

To use CMake, navigate to the top {\tt sw4} directory and run the following commands:
\begin{verbatim}
  > mkdir build
  > cd build
  > cmake [options] ..
  > make
  > make install
\end{verbatim}

The two dots after {\tt cmake [options]} are essential and instructs it to look in the parent
directory for the {\tt CMakeLists.txt} file.
 
The \verb+cmake+ command searches for the necessary libraries and other dependencies, then creates
makefiles that are appropriate for your system. You then run \verb+make+ to compiles and link
\emph{SW4} using these makefiles. For details about the exact commands being used in compilation,
run \texttt{make VERBOSE=1}.  Once SW4 has been successfully built, you will see the
``SW4 Lives!'' banner on the screen.

NOTE: If you want to rebuild \verb+sw4+ with a new set of options, you can force \verb+cmake+ to start
from scratch by removing the file \verb+CMakeCache.txt+ in the \verb+build+ directory. Another way
is to remove all files in the \verb+build+ directory.

\subsection{CMake Options}
CMake provides several options to allow customized configuration of \emph{SW4}.  To use any option,
add \texttt{-D\textless option\textgreater=\textless value\textgreater} to the options in the
\texttt{cmake} command.  For example, when building SW4 with the PROJ library:
%
\begin{verbatim}
  > export 
  > cmake -DUSE_PROJ=ON -DPROJ_DIR=${SW4ROOT}/install/lib64/cmake/proj ..
\end{verbatim}
%
A list of options is shown in the table below.
%
\begin{center}
\begin{tabular}{|l|c|p{0.55\textwidth}|}
\hline
Option & Default & Details \\
\hline
%
USE\_PROJ & OFF & Enable PROJ library. \\ \hline
PROJ\_DIR & (none) & The path to the PROJ installation. \\ \hline
%
USE\_HDF5 & OFF & Enable HDF5 library. \\ \hline
HDF5\_DIR & (none) & The path to the HDF5 installation. \\ \hline
%
USE\_ZFP & OFF & Enable ZFP compression. \\ \hline
ZFP\_DIR & (none) & The path to the ZFP installation. \\ \hline
H5Z\_ZFP\_DIR & (none) & The path to the H5Z-ZFP installation. \\ \hline
CMAKE\_BUILD\_TYPE & Release & The type of build to setup. Can be either \texttt{Debug},
\texttt{Release}, or \texttt{RelWithDebInfo}.  This affects the type of optimization and debug flags
used in compiling \emph{SW4}. \\ \hline
%
TESTING\_LEVEL & 0 & Specifies the testing level for automated tests.  Level 0 corresponds to tests
that run in roughly a minute or less (7 total), level 1 to tests that run in roughly 10 minutes or
less (13 total) and level 2 to tests that may require up to an hour or more (17 total). \\ \hline
%
MPI\_NUM\_TEST\_PROCS & 4 & Number of MPI processes to use in tests. Generally using more processes
will result in the tests finishing faster, but there is no point exceeding the number of available
cores on your system.  We strongly recommend at least 8 processes if TESTING\_LEVEL is 1 or
higher.\\ \hline
%
% -DMPIEXEC=/usr/bin/srun
%
MPIEXEC & mpirun & UNIX command for running an MPI application.\\ \hline
% -DMPIEXEC_NUMPROC_FLAG=-n
MPIEXEC\_NUMPROC\_FLAG & -np & MPI command option for specifying the number of processes.\\ \hline 
% -DMPIEXEC_PREFLAGS=-ppdebug
MPIEXEC\_PREFLAGS & (none) & Extra MPI command option.\\ \hline 
\end{tabular}
\end{center}

\paragraph{Modifying the MPI execution commands.}
By default, \verb+mpirun+ is used to start parallel runs when you do \verb+make test+.  However, on
Livermore computing (LC) machines the command for running MPI programs is \verb+srun+, not
\verb+mpirun+. Also, the flag for specifying the number of processors is different, and you must
give an additional flag for running interactive jobs on the debug partition. For example, you would
say
\begin{verbatim}
srun -ppdebug -n 128 sw4 inputfile.in
\end{verbatim}
to run on the debug partition using 128 cores. To modify the default MPI execution program and
other runtime parameters, the variables MPIEXEC, MPIEXEC\_NUMPROC\_FLAG, and
MPIEXEC\_PREFLAGS can be set as in the following example:
\begin{verbatim}
cmake -DTESTING_LEVEL=2 -DMPI_NUM_TEST_PROCS=128 -DMPIEXEC=srun \
      -DMPIEXEC_NUMPROC_FLAG=-n -DMPIEXEC_PREFLAGS=-ppdebug ..
\end{verbatim}

After the PROJ and HDF5 libraries have been installed (see next section), you need to tell
\verb+cmake+ where to find them. On the LC-machines, all three libraries are currently installed under
\verb+/usr/apps/wpp+, and you can use the following command options to configure \verb+sw4+:
\begin{verbatim}
cmake -DTESTING_LEVEL=2 -DMPI_NUM_TEST_PROCS=36 -DMPIEXEC=srun \
      -DMPIEXEC_NUMPROC_FLAG=-n -DMPIEXEC_PREFLAGS=-ppdebug \
      -DUSE_HDF5=ON -DPROJ_DIR=/usr/apps/wpp \
      -DUSE_PROJ=ON -DHDF5_DIR=/usr/apps/wpp ..
\end{verbatim}
To verify that \verb+cmake+ actually found the libraries, pay attention to the
following lines of the output from the \verb+cmake+ command:
\begin{verbatim}
...
-- Found PROJ: /usr/apps/wpp/lib/libproj.so  
...
\end{verbatim}

Sometimes CMake doesn't pick up the correct compiler. Say, for example that the C++ compiler on your
system is called {\tt mpicxx} and the Fortran compiler is {\tt mpiifort}. You can tell {\tt cmake}
to use those compilers by setting the following envoronment variables {\em before} running cmake
(assuming a {\tt csh} shell),
\begin{verbatim}
  > setenv CXX mpicxx
  > setenv FC mpiifort
\end{verbatim}

\subsection{CTest}\label{cha:ctest-sw4}
The \emph{SW4} CMake configuration includes several test cases that are used to verify the correctness
of the \emph{SW4} installation.  Each test consists of two parts. First it runs a case using an input
file in the \verb+pytest+ directory. Secondly, it checks that the results are within
a reasonable error tolerance from previously recorded results.

To run the tests, use either the command \texttt{make test} or \texttt{ctest} as follows:
%
\begin{verbatim}
build > ctest
Test project /Users/petersson1/src/sw4-cig/build
      Start  1: Run_twilight/flat-twi-1
 1/24 Test  #1: Run_twilight/flat-twi-1 .....................   Passed    0.49 sec
      Start  2: Check_Result_twilight/flat-twi-1
 2/24 Test  #2: Check_Result_twilight/flat-twi-1 ............   Passed    0.03 sec
      Start  3: Run_twilight/flat-twi-2
...
      Start 23: Run_pointsource/pointsource-sg-1
23/24 Test #23: Run_pointsource/pointsource-sg-1 ............   Passed   89.56 sec
      Start 24: Check_Result_pointsource/pointsource-sg-1
24/24 Test #24: Check_Result_pointsource/pointsource-sg-1 ...   Passed    0.03 sec

100\% tests passed, 0 tests failed out of 24

Total Test time (real) = 230.91 sec
\end{verbatim}

You can run tests selectively using \texttt{ctest -R \textless regex\textgreater}, for example:
%
\begin{verbatim}
build > ctest -R meshrefine
Test project /Users/petersson1/src/sw4-cig/build
    Start 15: Run_meshrefine/refine-el-1
1/6 Test #15: Run_meshrefine/refine-el-1 .................   Passed   25.61 sec
    Start 16: Check_Result_meshrefine/refine-el-1
2/6 Test #16: Check_Result_meshrefine/refine-el-1 ........   Passed    0.03 sec
    Start 17: Run_meshrefine/refine-att-1
3/6 Test #17: Run_meshrefine/refine-att-1 ................   Passed   22.00 sec
    Start 18: Check_Result_meshrefine/refine-att-1
4/6 Test #18: Check_Result_meshrefine/refine-att-1 .......   Passed    0.03 sec
    Start 19: Run_meshrefine/refine-att-2nd-1
5/6 Test #19: Run_meshrefine/refine-att-2nd-1 ............   Passed   17.63 sec
    Start 20: Check_Result_meshrefine/refine-att-2nd-1
6/6 Test #20: Check_Result_meshrefine/refine-att-2nd-1 ...   Passed    0.03 sec

100% tests passed, 0 tests failed out of 6

Total Test time (real) =  65.35 sec
\end{verbatim}

If a test fails you can check the details in the output log at
\texttt{Testing/Temporary/LastTest.log}.

\section{Installing the PROJ, HDF5, and ZFP}\label{sec:proj}
\index{installation!proj} \index{installation!hdf5}

If you are interested in using the advanced geographical mapping options of the {\tt grid} command, the {\tt sfile}, or the {\tt gmg} command, you need to install the PROJ package. For {\tt sfile}, {\tt ssioutput},  and {\tt gmg}, the HDF5 library is also required. To use the {\tt ssioutput} command with ZFP compression, both ZFP library and the ZFP HDF5 filter library (H5Z-ZFP) are required.  

The following instructions describe how to install the two packages. For simplicity all packages
are installed under the same top directory. If you are using {\tt cmake}, you may optionally put the
PROJ and HDF5 packages in a separate directory. In the following we shall assume that all packages are
installed under the same top directory, and that you assign the name of that directory to the
environment variable \verb+SW4ROOT+. When you are finished installing the packages, the
corresponding include and library files should be in the sub-directories \verb+${SW4ROOT}/include+
and \verb+${SW4ROOT}/lib+, respectively.

\subsection{PROJ}
The PROJ library requires SQLite, if your system does not have it installed already, it can be compiled with the following steps:
\begin{verbatim}
  # Download sqlite from https://www.sqlite.org/download.html (sqlite-autoconf-x)
  > tar xf sqlite-autoconf-x.tar.gz
  > cd sqlite-autoconf-x/
  > ./configure --prefix=${SW4ROOT}
  > make
  > make install
\end{verbatim}

The PROJ library (version 9 and later is recommended) can be installed by the following steps:

\begin{verbatim}
  # Download PROJ from https://proj.org/download.html
  > tar xf proj-x.x.x.tar.gz
  > cd proj-x.x.x
  > mkdir build
  > cd build
  > cmake -DBUILD_APPS=OFF -DCMAKE_INSTALL_PREFIX=${SW4ROOT}
-DSQLITE3_INCLUDE_DIR=${SW4ROOT}/include
-DSQLITE3_LIBRARY=${SW4ROOT}/lib/libsqlite3.so  ..

  # Note that the two -DSQLITE3 flags are needed if you compiled SQLite yourself.
  > make
  > make install
\end{verbatim}

\subsection{HDF5}
Installing HDF5 (version 1.12 and later is recommended) can be done with the following steps:
\begin{verbatim}
  # Download HDF5 from https://www.hdfgroup.org/downloads/hdf5/source-code
  > tar xf hdf5-1.xx.x.tar.gz
  > cd hdf5-1.xx.x
  > mkdir build
  > cd build
  > cmake -DHDF5_ENABLE_PARALLEL=ON -DCMAKE_INSTALL_PREFIX=${SW4ROOT}
  > make
  > make install
\end{verbatim}

\subsection{ZFP and H5Z-ZFP}
We recommend to use use ZFP and H5Z-ZFP's latest stable release version. 

Installing ZFP can be done with the following steps:
\begin{verbatim}
  # Download ZFP from https://github.com/LLNL/zfp/releases
  > tar xf zfp-x.x.x.tar.gz
  > cd zfp-x.x.x
  > mkdir build
  > cd build
  > cmake -DZFP_BIT_STREAM_WORD_SIZE=8 -DCMAKE_INSTALL_PREFIX=${SW4ROOT} ..
  > make
  > make install
\end{verbatim}

Installing H5Z-ZFP can be done with the following steps:
\begin{verbatim}
  # Download H5Z-ZFP from https://github.com/LLNL/H5Z-ZFP/releases
  > tar xf vx.x.x.tar.gz
  > cd H5Z-ZFP-x.x.x
  > mkdir build
  > cd build
  > export HDF_DIR=/path/to/hdf5/install
  > export ZFP_DIR=/path/to/zfp/install
  > cmake -DCMAKE_INSTALL_PREFIX=${SW4ROOT} ..
  > make
  > make install
\end{verbatim}

To verify that the libraries have been installed properly, you should go to the \verb+SW4ROOT+
directory and list the {\tt lib} sub-directory (\verb+cd ${SW4ROOT}; ls lib lib64+). You should see the
following files (on Mac OSX machines, the .so extension is replaced by .dylib ):
\begin{verbatim}
  > cd ${SW4ROOT}
  > ls lib lib64
  > lib
  > libhdf5.so ...
  > lib64
  > libproj.so ...
\end{verbatim}

Furthermore, if you list the include sub-directory, you should see include files such as 
\begin{verbatim}
  > cd ${SW4ROOT}
  > ls include
  > proj.h hdf5.h ...
\end{verbatim}

\subsection{SW4 with PROJ, HDF5, and ZFP}
To build SW4 with all the libraries can be done using CMake with the following commands:

\begin{verbatim}
  > export PROJ_ROOT=/path/to/proj/install
  > export HDF5_ROOT=/path/to/hdf5/install
  > export ZFP_ROOT=/path/to/zfp/install
  > export H5Z_ZFP_ROOT=/path/to/h5z_zfp/install
  > cmake -DUSE_HDF5=ON -DUSE_PROJ=ON -DUSE_ZFP=ON ..
  > make
\end{verbatim}



\section{Disclaimer} 
This document was prepared as an account of work sponsored by an agency of the United States
government. Neither the United States government nor Lawrence Livermore National Security, LLC, nor
any of their employees makes any warranty, expressed or implied, or assumes any legal liability or
responsibility for the accuracy, completeness, or usefulness of any information, apparatus, product,
or process disclosed, or represents that its use would not infringe privately owned
rights. Reference herein to any specific commercial product, process, or service by trade name,
trademark, manufacturer, or otherwise does not necessarily constitute or imply its endorsement,
recommendation, or favoring by the United States government or Lawrence Livermore National Security,
LLC. The views and opinions of authors expressed herein do not necessarily state or reflect those of
the United States government or Lawrence Livermore National Security, LLC, and shall not be used for
advertising or product endorsement purposes. 

\bibliographystyle{plain}

\bibliography{sw4-refs} 

\end{document}
