%\documentclass[12pt,reqno]{amsart}
\documentclass[twoside,11pt]{article}
\usepackage{amsmath,amssymb,amsthm,amsfonts}
\topmargin=-0.5in \oddsidemargin3mm \evensidemargin3mm
\textheight225mm \textwidth160mm
\usepackage{xcolor}
 \usepackage{natbib}
\usepackage{graphicx}
\linespread{1.125}
\begin{document}

\begin{center}
{\Large \textbf{Responses to Referee Report (Reviewer \#2)}}
\end{center}

\noindent{Dear editor and referees of the SIAM Journal on Scientific Computing},

We would like to thank you for the time and help in handling and carefully reading our submission and providing encouraging and constructive comments.  We have revised our paper accordingly and improved its context in several courses. 

We have added a numerical example of the benchmark problem LOH.1 in Section 5.5.

We have also carefully proofread the manuscript, polished its writing and corrected the typos and grammatical errors.  Here are the comments from reviewer \#2 in bold followed by our responses.

\begin{enumerate}
\item \textbf{The first major concern of this reviewer is the rather complicated notations adopted by the authors. I think the authors are trying to use a much more compact notations. This does not really work for (me) the reader. Using variables and parameters with multiple subscripts and superscripts, at the same time, makes the presentation cumbersome, and stands in the way of the reader. For this reason I was not able to read all the analysis, but I think the analysis is correct since it based on the author’s previous papers} \cite{petersson2009stable,wang2019fourth,petersson2015wave}.\\
 We have gone through the manuscript and simplified notations in some formulas and proofs.

\item  \textbf{The use of method of manufactured solution might be useful to check the correctness of code but it is not sufficient to verify the accuracy of method in a seemingly realistic setting. The manufactured solutions may not exhibit the properties of the real solutions, for example large gradients, in the presence of discontinuous material properties. As the authors are aware, there are community developed benchmark problems, with analytical solutions, to test the accuracy of simulation codes for seismic wave propagation problems. In particular the LOH1 benchmark problem of SCEC (Southern California Earthquake Center) see also http://www.sismowine.org/model.html with discontinuous wavespeeds and density should be a good candidate for this test}: We have the test for LOH.1 problem in Section 5.5. We observe very good agreement with the exact solution. In addition, the computational time of solving the linear system for interface coupling is only $4\%$ of the time-stepping procedure. In comparison, it is $8\%$ with the traditional SBP-GP method.

\item  \textbf{The authors should explicitly mention that the scheme is semi-implicit, since it involves numerical solutions of linear algebraic systems to impose the interface/boundary conditions. The solutions of systems of linear algebraic problem will certainly be nontrivial for large scale simulation of seismic waves in realistic 3D settings. The authors should explain how the implicit part of the scheme will affect parallel scalability which is necessary for large scale simulation, and in particular in 3D. For several reasons, purely explicit schemes seem to be preferable for large scale wave propagation problems. How will the semi-implicit method developed here compare with purely explicit schemes such as those developed in} \cite{virta2014acoustic,duru2014stable}: We have added in the introduction that the ghost point values are obtained by solving a linear system of equations. In the LOH.1 test problem, we observe that the method implemented in SW4 has excellent parallel scalability. Considering its better CFL property comparing with the purely explicit methods by SAT, our method could be more efficient. 

\item  \textbf{To me, the theoretical results presented in the paper are direct extensions of} \cite{wang2019fourth} \textbf{to 3D and/or} \cite{petersson2009stable} \textbf{to 4th order accuracy. The theory in the paper is not really novel. And since the method has been implemented in SW4, which is a scalable open source production software: Are there problems that were difficult/impossible for SW4 that are now accessible using the new method? The numerical examples presented are too simplistic, in my opinion. I expected to see a much more interesting simulations of scenarios closer to realistic settings. This will help make the case for} \cite{wang2019fourth}\textbf{, and its extensions to 3D}: For the original method in SW4, it cannot deal with curved mesh-refinement interfaces but this can be done with our method. In addition, the new method using ghost points only from the coarse domain improves computational efficiency as we have observed from running the LOH.1 benchmark problem in Section 5.5.

\item \textbf{In the introduction the unstaggered grid methods} \cite{kozdon2013simulation,duru2016dynamic} \textbf{are worth mentioning}.\\
We have added these two references in the introduction section.

\end{enumerate}

Again, we appreciate you all for the time in handling and scrutinizing our manuscript, and thanks for considering our submission to this journal.

Sincerely, 

All authors

\bibliography{revier1.bib}
\bibliographystyle{plain}

\end{document}
